Converging High-Throughput and High-Performance Computing Paradigms: \\ A Case Study of Using Abstractions to Bridge the Divide

### 2017-03-16 PandaEvolution.tex
Iteration 4 and 5 adding Sergey's data

\subsection{Use of Supercomputers with PanDA}

@@

% -----------------------------------------------------------------------------
\subsection{Interfacing PanDA with Titan}
\label{ssec:panda_titan}

@@

\mtnote{Does one PanDA Broker support the submission and monitoring of one job
at a time? Or does it support the concurrent submission and monitoring of
multiple jobs (assuming enough backfill availability)?}

Once deployed, every PanDA Broker supports the execution of MC detector
simulation on Titan in 9 steps: (i) PanDA Broker queries Titan's Moab scheduler
about currently unused worker nodes; (ii) the PanDA Broker queries the Job
Dispatcher module of the PanDA server for jobs that have been bound to Titan by
JEDI; (iii)  the PanDA Broker requests to the Data Service module of the PanDA
Server what input data these jobs require and transfer them to the DTN via the
ATLAS DDM service; (iv) the PanDA Broker creates one or more PBS MPI script,
wrapping enough jobs' payload to fit backfill availability; (v) the PanDA Broker
submits these MPI jobs to the Titan's PBS batch system via RADICAL-SAGA; (vi)
the MPI jobs waits a minimal amount of time into the queue until they are
scheduled for execution; (vii) upon execution on the worker nodes, each MPI
script creates a ramdisk and sets up the specific AthenaMP configuration
required by the wrapped payload; (viii) the payload is executed into dedicated
processes, 16 for each work node, 1 for each available core; (ix) upon
completion of an MPI job, the PanDA Broker locates the produced output files on
the shared filesystem and transfer them to BNL ATLAS Tier 1 computing centre.

>>>>

% -----------------------------------------------------------------------------
\subsection{Use of Supercomputers with PanDA}
\label{ssec:panda-supercomputers}

@@

% -----------------------------------------------------------------------------
\subsection{Interfacing PanDA with Titan}
\label{ssec:panda-titan}

@@

Once deployed, every PanDA Broker supports the execution of MC detector
simulation on Titan in 8 steps: (i) PanDA Broker queries Titan's Moab scheduler
about current backfill availability; (ii) the PanDA Broker queries the Job
Dispatcher module of the PanDA server for jobs that have been bound to Titan by
JEDI; (iii)  the PanDA Broker requests to the Data Service module of the PanDA
Server what input data these jobs require and transfer the files to the DTN via
the ATLAS DDM service; (iv) the PanDA Broker creates a PBS MPI script, wrapping
enough jobs' payload to fit backfill availability; (v) the PanDA Broker submits
the PBS MPI script to the Titan's PBS batch system via RADICAL-SAGA; (vi) upon
execution on the worker node(s), the MPI script creates a ramdisk and configure
AthenaMP to execute the wrapped payload; (vii) the payload is executed into
dedicated processes, 16 for each work node, 1 for each available core; (viii)
upon completion of the MPI script, the PanDA Broker locates the output on the
shared NFS filesystem and transfer it to BNL ATLAS Tier 1 computing centre.

### 2017-03-16 overall.tex
Iteration 4 and 5 adding Sergey's data

% -----------------------------------------------------------------------------
\subsection{Efficiency and Scalability of PanDA Broker on Titan}
\label{ssec:panda_titan}

@@

All the detector simulations executed on Titan process 100 events. This number
has been chosen by observing the average walltime of backfill resources: too
many events would \ldots\mtnote{I need to speak further to Sergey to clarify
this.}.

Usually, detector simulations performed on the Grid process a higher number of
events than those on Titan. This explains the difference between the percentage
of detector simulations (9\%) and of events computed by those simulations
(3.9\%) on Titan. PanDA Event Service~\cite{panda-event-service} will contribute
to address this limitation: each PanDA Broker will be able to pull a variable
number of events for each simulation, depending on the available walltime. This
functionality is currently being implemented.

Given the current design and implementation, the utilization of a larger portion
of Titan's backfill availability depends on the the number of concurrent PanDA
Brokers instantiated on the DTNs.
% The number of concurrent detector simulation jobs that PanDA Broker can manage
% depends on the DTNs' resources. Each PanDA Broker can execute 1 MPI PBS job on
% up to 300 work nodes for a total of 4,800 cores. Scaling above this threshold
% requires instantiating multiple PanDA Brokers on the same DTN and, once the
% DTN resources are saturated, on multiple DTNs.
In principle, this is not a problem on Titan as multiple DTNs are supposed to be
available and periodically upgraded. Assuming the efficiency reached on February
as reference, using all the available backfill resources on Titan would require
around 60 concurrent PanDA Brokers. Further testing and better monitoring
capabilities are required to establish whether the current DTN capabilities can
support this load, especially when considering the increased volume of files
staged in and out the DTNs.

\mtnote{Extend/change previous paragraph explaining also the problem of the
`spike'/`wave' of backfill availability. The problem without enough brokers to
cope with variable amounts of backfill availability is that we may exhaust all
20 brokers to use a portion of backfill smaller than the one that becomes
available when no more brokers are available.}

@@

At runtime, detector simulation jobs showed performance degradation when all the
16 cores of Titan's work nodes were used. As shows in
Figure~\ref{}\mtnote{Sergery: do you have a diagram showing the difference in
performance between using 16 and 8 cores for a detector simulation job with 100
events?}, the mean runtime for 8 concurrent jobs computing 100 events is XXXs,
with each job executed in 8 distinct processes, one per core. The mean runtime
for 16 concurrent jobs computing 100 events is instead XXXs. This can be
explained by looking at the cache architecture of the AMD Opteron 6274 CPU. In
this processor, every two cores share the level 1 instruction cache, impacting
the performance of \ldots\mtnote{Do we kow what instruction set is affected?}
computation.

>>>>

% -----------------------------------------------------------------------------
\subsection{Utilization and Efficiency of PanDA Broker on Titan}
\label{ssec:panda_titan}

@@

All the detector simulations executed on Titan process 100 events. This number
of events is consistent with the physics of the use case and with the average
duration of backfill availability. The duration of a detector simulation is a
function of the number of events: the more events, the longer the simulation.
Not all events take the same time to be simulated:
Figure~\ref{fig:100event-distrib} shows the distribution of that duration for
XXXXX events simulated on Titan. When compared to the distribution of backfill
availability on Titan in Figure~\ref{fig:backfill-distrib}, 100 events appear to
be a good compromise.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
\caption{Distribution of the time taken by a Geant4 detector simulation to simulate one event. 3000 Events; 30 Titan work nodes; 16 simulations per node; 100 events per node. Average time per event \~14 min. Broad distribution from \~1 to \~40 minutes.}
\label{fig:100event-distrib}
\end{figure}

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
\caption{\mtnote{Placeholder for the distribution of backfill walltime availability on Titan.}}
\label{fig:backfill-distrib}
\end{figure}

Usually, detector simulations performed on the Grid process a higher number of
events than those on Titan. This explains the difference between the percentage
of detector simulations (9\%) and of events computed by those simulations
(3.9\%) on Titan. PanDA Broker could fit the number of events to the walltime of
the backfill availability on the base of the distribution of
Figure~\ref{fig:backfill-distrib}. That specific number of event could then be
pulled from the PanDA Event service~\cite{panda-event-service} and given as
input to one or more simulations. Once packaged into a MPI PBS job, these
simulations would better fit backfill availability, increasing the efficiency of
PanDA Brokers.

The transition from a homogeneous to a heterogeneous number of events per
detector simulation has implications for the application layer. An even number
of events across simulations makes it easier to partition, track and package
events across simulations, especially when they are performed on both the Grid
and Titan. A homogeneous number of events also helps to keep the size and
duration of other stages of the ATLAS MC workflow (\S\ref{ssec:panda-titan})
more uniform. Further analysis is needed to evaluate the trade offs between
increased efficiency of resource utilization and the complexity that would be
introduced at the application layer.

Currently, each PanDA Broker creates, submits, and monitor a single MPI PBS
script at a time. This design is inherited from PanDA Pilot where a single
process is spawn at a time to execute the payload. As a consequence, the
utilization of a larger portion of Titan's backfill availability depends on the
the number of concurrent PanDA Brokers instantiated on the DTNs: When all the 20
PanDA Brokers have submitted a MPI PBS Job, further backfill availability cannot
be used.

Based on the current data, increasing PanDA Brokers efficiency on Titan would
require around 60 concurrent brokers. Further testing and better monitoring
capabilities are required to establish whether the DTN capabilities can support
this load, especially when considering the volume of files staged in and out the
DTNs. Alternatively, a design of a PanDA Broker capable of managing multiple MPI
scripts at a time is being evaluated.

% The number of concurrent detector simulation jobs that PanDA Broker can manage
% depends on the DTNs' resources. Each PanDA Broker can execute 1 MPI PBS job on
% up to 300 work nodes for a total of 4,800 cores. Scaling above this threshold
% requires instantiating multiple PanDA Brokers on the same DTN and, once the
% DTN resources are saturated, on multiple DTNs.

@@

The AMD Opteron 6274 CPU used on Titan has 16 cores, divided into 8 compute
units. Each compute units has 1 floating point (FP) scheduler shared between 2
cores. When using 16 cores for FP-intensive calculations, each pair of cores
competes for a single FP scheduler. This creates the overhead shown in
Figure~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
concurrent simulations computing 50 events is 10.8 minutes, while for 16
simulations is 14.25 minutes (consistent with
Figure~\ref{fig:100event-distrib}). The allocation policy based on work nodes
and not cores used, does not justify the use of 1/2 of the cores available.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/tx8_tx16_comparison.pdf}
\caption{Comparison between distributions of the time taken by a Geant4 detector simulation to simulate one event when placing 2 simulations (h1) or 1 simulation (h2) per CPU. 2 simulation use 16 cores per node, 1 simulation 8, 1 per compute unit. 50 Events; 1 Titan work nodes; 16 work threads per node; 100 events per node. Average time per event \~14 min. Broad distribution from \~1 to \~40 minutes.}
\label{fig:comparison-8-16cores}
\end{figure}
